% -----------------------------------------------------------------------
% future.tex: Future possible developments for Duchamp.
% -----------------------------------------------------------------------
% Copyright (C) 2006, Matthew Whiting, ATNF
%
% This program is free software; you can redistribute it and/or modify it
% under the terms of the GNU General Public License as published by the
% Free Software Foundation; either version 2 of the License, or (at your
% option) any later version.
%
% Duchamp is distributed in the hope that it will be useful, but WITHOUT
% ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
% FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License
% for more details.
%
% You should have received a copy of the GNU General Public License
% along with Duchamp; if not, write to the Free Software Foundation,
% Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307, USA
%
% Correspondence concerning Duchamp may be directed to:
%    Internet email: Matthew.Whiting [at] atnf.csiro.au
%    Postal address: Dr. Matthew Whiting
%                    Australia Telescope National Facility, CSIRO
%                    PO Box 76
%                    Epping NSW 1710
%                    AUSTRALIA
% -----------------------------------------------------------------------
\secA{Future developments}

Here are lists of planned improvements and a wish-list of
features that would be nice to include (but are not planned in the
immediate future). Let me know if there are items not on these lists,
or items on the list you would like prioritised.

Planned developments:
\begin{itemize}
\item Parallelisation of the code, to improve speed particularly on
multi-core machines.

\item Better determination of the noise characteristics of
  spectral-line cubes, including understanding how the noise is
  generated and developing a model for it. 
  
\item Include more source analysis. Examples could be: shape
  information; measurements of HI mass; more variety of measurements
  of velocity width and profile. 

\item Improved ability to reject interference, possibly on the
  spectral shape of features.

\item Ability to separate (de-blend) distinct sources that have been
  merged.
\end{itemize}

Wish-list:
\begin{itemize}
\item Incorporation of Swinburne's S2PLOT
\footnote{\href{http://astronomy.swin.edu.au/s2plot/}
{http://astronomy.swin.edu.au/s2plot/}} code for improved
visualisation. 
\item Link to lists of possible counterparts (\eg via NED/SIMBAD/other
  VO tools?). 

\item On-line web service interface, so a user can upload a cube and
  get back a source-list.

\item Embed \duchamp in a GUI, to move away from the text-based
  interaction.
\end{itemize}

\secA{Why ``\duchamp''?}

Well, it's important for a program to have a name, and the initial
working title of \emph{cubefind} was somewhat uninspiring. I wanted to
avoid the classic astronomical approach of designing a cute acronym,
and since it is designed to work on cubes, I looked at naming it after
a cubist. \emph{Picasso}, sadly, was already taken \citep{minchin99},
so I settled on naming it after Marcel Duchamp, another cubist, but
also one of the first artists to work with ``found objects''.

