\begin{thebibliography}{9}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{{\tt #1}}\fi

\bibitem[{Calabretta} and {Greisen}(2002)]{calabretta02}
M.R. {Calabretta} and E.W. {Greisen}.
\newblock {``Representations of celestial coordinates in FITS''}.
\newblock
\href{http://adsabs.harvard.edu/cgi-bin/nph-bib_query?bibcode=2002A%26A...395%.1077C&db_key=AST}%
{{\em A\&A}, 395:\penalty0 1077--1122, December 2002.}
%\newblock URL
.

\bibitem[{Greisen} and {Calabretta}(2002)]{greisen02}
E.W. {Greisen} and M.R. {Calabretta}.
\newblock {``Representations of world coordinates in FITS''}.
\newblock
\href{http://adsabs.harvard.edu/cgi-bin/nph-bib_query?bibcode=2002A%26A...395%.1061G&db_key=AST}%
{{\em A\&A}, 395:\penalty0 1061--1075, December 2002.}
%\newblock URL
%  \url{.

\bibitem[{Hanisch} et~al.(2001){Hanisch}, {Farris}, {Greisen}, {Pence},
  {Schlesinger}, {Teuben}, {Thompson}, and {Warnock}]{hanisch01}
R.J. {Hanisch}, A.~{Farris}, E.W. {Greisen}, W.D. {Pence}, B.M. {Schlesinger},
  P.J. {Teuben}, R.W. {Thompson}, and A.~{Warnock}.
\newblock {``Definition of the Flexible Image Transport System (FITS)''}.
\newblock 
\href{http://adsabs.harvard.edu/cgi-bin/nph-bib_query?bibcode=2001A%26A...376%
..359H&db_key=AST}%
{{\em A\&A}, 376:\penalty0 359--380, September 2001.}
%\newblock URL
%  \url.

\bibitem[{Hopkins} et~al.(2002){Hopkins}, {Miller}, {Connolly}, {Genovese},
  {Nichol}, and {Wasserman}]{hopkins02}
A.M. {Hopkins}, C.J. {Miller}, A.J. {Connolly}, C.~{Genovese}, R.C. {Nichol},
  and L.~{Wasserman}.
\newblock {``A New Source Detection Algorithm Using the False-Discovery
  Rate''}.
\newblock 
\href{http://adsabs.harvard.edu/cgi-bin/nph-bib_query?bibcode=2002AJ....123.1%
086H&db_key=AST}%
{{\em AJ}, 123:\penalty0 1086--1094, February 2002.}
%\newblock URL
%  \url.

\bibitem[Lutz(1980)]{lutz80}
R.K. Lutz.
\newblock {``An algorithm for the real time analysis of digitised images''}.
\newblock {\em The Computer Journal}, 23:\penalty0 262--269, 1980.

\bibitem[{Meyer} et~al.(2004)]{meyer04:trunc}
M.J. {Meyer} et~al.
\newblock {``The HIPASS catalogue - {\sc I}. Data presentation''}.
%\newblock 
\href{http://adsabs.harvard.edu/cgi-bin/nph-bib_query?bibcode=2004MNRAS.350.1%
195M&db_key=AST}%
{{\em MNRAS}, 350:\penalty0 1195--1209, June 2004.}
%\newblock URL
%  \url.

\bibitem[{Miller} et~al.(2001){Miller}, {Genovese}, {Nichol}, {Wasserman},
  {Connolly}, {Reichart}, {Hopkins}, {Schneider}, and {Moore}]{miller01}
C.J. {Miller}, C.~{Genovese}, R.C. {Nichol}, L.~{Wasserman}, A.~{Connolly},
  D.~{Reichart}, A.~{Hopkins}, J.~{Schneider}, and A.~{Moore}.
\newblock {``Controlling the False-Discovery Rate in Astrophysical Data
  Analysis''}.
\newblock 
\href{http://adsabs.harvard.edu/cgi-bin/nph-bib_query?bibcode=2001AJ....122.3%
492M&db_key=AST}%
{{\em AJ}, 122:\penalty0 3492--3505, December 2001.}
%\newblock URL
%  \url.

\bibitem[Minchin(1999)]{minchin99}
R.F. Minchin.
\newblock {``Finding the Bivariate Brightness Distribution of Galaxies from an
  HI Selected Sample''}.
\newblock {\em PASA}, 16:\penalty0 12--17, 1999.

\bibitem[Starck and Murtagh(2002)]{starck02:book}
J.-L. Starck and F.~Murtagh.
\newblock {\em {``Astronomical Image and Data Analysis''}}.
\newblock Springer, 2002.

\end{thebibliography}
